### 2012-10-07 Recomb/conclusion_RECOMB.tex
bib less space/conclusion

%!TEX root = main_RECOMB.tex
\section{Conclusion}
\label{sec:conclusion}

In this article we presented a new and efficient way of
exploring the mutational landscape of an RNA, benefiting from the
 conserved consensus secondary structures  information,
to identify and fix sequencing errors. We also showed a new
 pseudo energy model, taking into account the nearest-neighbour energy model 
and, to quantify geometrical differences,  the isostericity of all base pairs.

By combining those two approaches,  the 
mutational landscape exploration and the pseudo energy model,
 into \texttt{RNApyro}, we show that we can
correct randomly mutated 5s rRNA sequences. 
As presented in Sec.~\ref{sec:results},
we observe that the models
with larger weights on the
isostericity seems to hold a higher accuracy on the estimation of errors.
Importantly, the implementation is fast enough for practical applications.

We must point out that our approach can only detect point wise 
sequencing errors, and only if they are located in base pairs. It 
nonetheless supplements well existing tools by using previously discarded
information.