### 2012-10-07 Recomb/conclusion_RECOMB.tex

%!TEX root = main_RECOMB.tex
\section{Conclusion}
\label{sec:conclusion}

In this article we presented a new and efficient way of
exploring the mutational landscape of an RNA, benefiting from the
 conserved consensus secondary structures  information,
to identify and fix sequencing errors. We also showed a new
 pseudo energy model, taking into account the nearest-neighbour energy model 
and, to quantify geometrical differences,  the isostericity of all base pairs.

By combining those two approaches,  the 
mutational landscape exploration and the pseudo energy model,
 into \texttt{RNApyro}, we show that we can
correct randomly mutated 5s rRNA sequences. 
As presented in Sec.~\ref{sec:results},
we observe that the models
with larger weights on the
isostericity seems to hold a higher accuracy on the estimation of errors.
Importantly, the implementation is fast enough for practical applications.

We must recall that our approach is restricted to
 the correction of point-wise error in structured regions (i.e. base paired nucleotides).
 Nonetheless it supplements well existing tools by using previously discarded
information holding, as shown, a strong signal.