### 2012-12-18 ISMB/conclusion_ISMB.tex
started ISMB folder

%!TEX root = main_RECOMB.tex
\section{Conclusion}
\label{sec:conclusion}