### 2013-01-04 ISMB/conclusion_ISMB.tex
fixed ISMB headers

%!TEX root = main_ISMB.tex
\section{Conclusion}
\label{sec:conclusion}