### 2013-01-04 ISMB/introduction_ISMB.tex

%!TEX root = main_ISMB.tex
\section{Introduction}
\label{sec:introduction}