### 2013-01-07 ISMB/methods_ISMB.tex
Method... interrupted

%!TEX root = main_ISMB.tex
\section{Methods}
\label{sec:methods}

We introduce a probabilistic model for the design of RNA sequences with a specific $\Gb+\Cb$ content and folding into a predefined secondary structure.
For the sake of simplicity, we choose to base this proof-of-concept implementation on a simplified free-energy function $\ES(\cdot)$, which only considers the contributions of 
stacked canonical base-pairs.

We show how a modification of the dynamic programming scheme introduced in the previously introduced \RNAmutants allow for the sampling of good and diverse design candidates in linear time and space complexities.

%To that purpose, a Boltzmann weighted distribution is used, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment. In this model, the probability that the nucleotide at a given position needs to be mutated (i.e. corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990}.
%
%\subsection{Probabilistic model}
%Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
%then any sequence $s$ is assigned a probability proportional to its Boltzmann factor
%\begin{align*}
%  \mathcal{B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
%\end{align*}
%where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
%are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.
%
%\subsubsection{Energy contribution}
\subsection{Definitions}

%\subsubsection{Profile}
%At every position, the nucleotide distribution is the same. 
%It is first initialized equiprobably. A bisection scheme is then applied until the sampled sequences $\Cb+\Gb$ content is as desired. The probability
%of a given nucleotide $a\in\B$ will be given by $\Prob(a)$ at every 
%position.

\subsubsection{Secondary structure}
Todo...

\subsubsection{Energy model}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using individual values from the Turner 2004 model (retrieved from the NNDB~\cite{Turner2010}). Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. not in $\{\Gb\Ub,\Ub\Gb,\Cb\Gb,\Gb\Cb, \Ab\Ub\text{ or }\Ub\Ab\}$). This latter parameter allows one to choose whether to simply penalize invalid base pairs ($\beta>0$), or forbid them altogether ($\beta = \infty$).

\subsection{\GCContent weighted pseudo-Boltzmann ensemble and distribution}

In order to counterbalance the documented tendency of sampling methods to generate \Gb\Cb-rich sequences~\cite{Levin:2012kx}, let us introduce a parameter $x\in\mathbb{R}^+$, whose value will influence the \GCContent of generated sequences. Let $\Target$ be a targeted secondary structure and $\gc(s)$ be the number of occurrences of \Gb and \Cb in an RNA sequence $s$, then let us define the pseudo-Boltzmann factor $\B_{x}(s,S)$ of a sequence $s$ such that
\begin{equation}
\B_{x, \Target}(s) = e^{\frac{-\ES(s,\Target)}{RT}}\cdot x^{\gc(s)}
\label{def:genBoltz}
\end{equation}
where $R$ is the Boltzmann constant and $T$ the temperature in Kelvin.

Summing the pseudo-Boltzmann factor over all possible sequences of a given length $|\Target|$ gives the pseudo-partition function $\mathcal{Z}_{\Target,x}$, from which one defines the pseudo-Boltzmann probability $\Prob_{x,\Target}(s)$ of each sequence $s$, respectively such that 
\begin{align}\mathcal{Z}_{x,\Target} &= \sum_{\substack{|s|=|\Target|}}\B_{x,\Target}(s)& \text{and}&&
\Prob_{x,\Target}(s) &= \frac{\B_{x,\Target}(s)}{\mathcal{Z}_{x,\Target}}.\label{def:distribution}\end{align}

It was established~\cite{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the \GCContent distribution observed in sequences sampled with respect to the above distribution converges to a normal law of expectation in $\mu\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, it can easily be shown that the expected \GCContent is a continuous and strictly increasing monotonic  function of $x$, whose limits are $0$ when $x=0$ and $n$ when $x\to +\infty$.

Consequently, for any targeted \GCContent $gc\in[0\%,100\%]$, there exists a unique value $x_{gc}$ such that generated sequences feature, on the average, the right \GCContent. Such a value can be determined using any numerical recipe, and a simple bisection is used in our implementation. Furthermore, the concentration of the distribution, as asserted by the limited standard deviation, implies a growth in $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. constant) for the expected number of attempts before generating a sequence having the exact targeted \GCContent  (resp. within $[gc-\kappa,gc+\kappa]$, where $\kappa\in\Theta(1/\sqrt n)$).

\subsection{Sampling the pseudo-Boltzmann ensemble}

Let us now describe a linear-time algorithm to sample sequences at random in the pseudo-Boltzmann distribution. This algorithm follows the general principles of the recursive approach to random generation~\cite{Wilf1977}, pioneered in the context of RNA by the \SFold algorithm~\cite{Ding2003}. The algorithm starts by precomputing the partition function restricted to each valid interval, and then performs a series of stochastic backtracks, using precomputed values to decide on the probability of each alternative.

\subsubsection{Precomputation}
Let us describe a dynamic programming algorithm to compute $\Z{i,j}{a,b}$, the pseudo-partition function restricted to an interval $[i,j]$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively.

Since the only sequence over an empty interval is the empty sequence, having energy $0$, one has
\begin{equation}
	\forall i \in [0,n-1]:\, \Z{i+1,i}{a,b}=1.
	\label{eq:Z_in}
\end{equation}

The recursion consists in four terms, depending on the base-pairing status and context of $(i,j)$:
\begin{equation*}
	\Z{i,j}{a,b}:=\left\{
  \begin{array}{ll}
  		\displaystyle
      \sum_{a'\in \B}  
      x^{\gc(a')}
      \cdot\Z{i+1,j}{a',b} &\begin{array}{@{}l@{}}\text{If }\Target_{i}=-1\\ \text{\relsize{-1}Position $i$ unpaired,}\end{array}\\
      \displaystyle
      \sum_{a',b'\in \B^2}
			 x^{\gc(a'.b')}
			 \cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 \cdot \Z{i+1,j-1}{a',b'}&
\begin{array}{@{}l@{}}\text{Elif }\Target_i=j \land \Target_{i-1}=j+1\\ \text{\relsize{-1}$(i,j)$ base-paired, stacked onto $(i$-$1,j$+$1)$,}\end{array}
			 \\
			 \displaystyle
      \sum_{a',b'\in \B^2}
      x^{\gc(a'.b')}
			\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
      \cdot\Z{i+1,k-1}{a',b'}
      \cdot\Z{k+1,j}{b',b} & 
\begin{array}{@{}l@{}}\text{Elif }\Target_i=k \land i < k \leq j\\
\text{\relsize{-1} Junction or bulge $\Rightarrow$ split, no stacking.}
\end{array}
\\
      0 &\text{Otherwise}
	\end{array}\right.
\label{eq:Z_rec}
\end{equation*}
It is worth mentioning that, despite the appearances, this recurrence can be computed in linear $\Theta(|\Target|)$ time and space. Indeed, one can show that position $j$ is entirely determined by $i$ within subsequent recursive calls.

More precisely, let us consider the following predicate: {\em Let $(i',j')$ be the base-pair such that $i'<i<j'$ while minimizing $i-i'$, then TODO!!!!

$j$ is the position preceding the end of the closest base-pair nesting position $i$ if such a base-pair exists, or the last position in the sequence otherwise.} Obviously, the predicate holds for $[0,n-1]$. Moreover, if $i$ is unpaired, then $i+1$ and $i$

%The cases can be broken down as follows:
%\begin{description}
%\item[$S_{i}=-1$:] If the nucleotide at position $i$ is unpaired, then 
%any sequence consists in a nucleotide $a'$ at position  $i$, 
%followed by a sequence over $[i+1,j]$ and having flanking nucleotides $a'$ and $b$.
%\item[$S_i=j$ and $S_{i-1}=j+1$:] Any sequence generated in $[i,j]$ consists of two nucleotides $a'$ and $b'$, flanking a sequence over $[i+1,j-1]$.
%Since positions $i$ and $i-1$ are paired with $j$ and $j+1$ respectively, 
%then a stacking energy contribution is added. 
%\item[$S_i=k$ and $i<k \leq j$:] If position $i$ is paired and not involved in a stacking, then the 
%only term contributing directly to the energy is the nucleotide distribution $\Prob(\cdot)$. 
%Any sequence on $[i,j]$ consists of two nucleotide $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence over interval $[i+1,k-1]$ and preceding a (possibly empty) sequence interval $[k+1,j]$.
%\item[Else:] In any other case, we are in a derivation of the SCFG that does not correspond to the secondary structure $S$, and we return $0$.
%\end{description}

\subsubsection{Stochastic backtrack}
Once the partition functions have been computed, a stochastic backtrack starts from the complete interval $[0,|\Target|-1]$ and, at each step, chooses a suitable assignment for one or several positions, as described in Algorithm~\ref{alg:back}. 
\begin{algorithm}[t]
	\SetAlgoLined
\SetKwFunction{Backtrack}{Backtrack}
\SetKwFunction{Random}{Random}
	rand $\leftarrow$ \Random$\left(\Z{i,j}{a,b}\right)$\;
	k $\leftarrow S_i$\;
	\If(\tcp*[f]{Unpaired}){$k = -1$}{
		\For{$a'\in\B$}{
			rand $\leftarrow$ rand $- x^{\gc(a')}\cdot \Z{i+1,j}{a',b}$\;
			\lIf{$\text{rand}<0$}\Return{$a'.\Backtrack\left((i+1,j),(a',b),S\right)$}\;
		}
	}
	\Else{
		\If(\tcp*[f]{Stacking pairs}){$k=j \land S_{i-1}=j+1$}{
			\For{$(a',b')\in\B\times\B$}{
				rand $\leftarrow$ rand $-
			 	x^{\gc(a'.b')}
			 	\cdot e^{\frac{-\ES^{\beta}_{ab \to a'b'}}{RT}}
			 	\cdot \Z{i+1,j-1}{a',b'}	$\;
				\If{$\text{rand}<0$}{
					\Return{$a'.\Backtrack\left((i+1,j-1),(a',b'),S\right). b'$}\;		
				}
			}
		}
		\Else{
			\If(\tcp*[f]{Paired, not stacked}){$\Target_i=k \land i < k \leq j$}{
				\For{$(a',b')\in\B\times\B$}{
					rand $\leftarrow$ rand $-	
       x^{\gc(a'.b')}
					\cdot e^{\frac{-\ES^{\beta}_{\varnothing\to a'b'}}{RT}}
  	    	\cdot\Z{i+1,k-1}{a',b'}
	    	  \cdot\Z{k+1,j}{b',b}$\;	
 					\If{$\text{rand}<0$}{
						\Return{$a'
						.\Backtrack\left((i+1,k-1),(a',b'),S\right)
						.b'
						.\Backtrack\left((k+1,j),(b',b),S\right)
						$}\;	
					}
				}
		 }
    \Else {\Return{$\varepsilon$};	\tcp*[f]{Empty sequence}}
		}
	}
\caption{\protect\Backtrack$\left((i,j),(a,b),S\right)$\label{alg:back}}
\end{algorithm}

\subsection{Overall complexity}

%\subsubsection{Outside}	
%\begin{figure}
%\resizebox{\textwidth}{!}{\input{FigDPOutside.pgf}}
%\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
%therefore its efficient implementation requires an implementation of the inside computation.}
%\end{figure}
%The \emph{Outside} function, $\mathcal Y$, is the partition function considering only the 
%contributions of subsequences $[0,i]\cup[j,n-1]$ over the mutants of $s$ having exactly $m$ mutations between $[0,i]\cup[j,n-1]$ and whose nucleotide at position $i+1$ is $a$ 
%(resp. in position $j-1$ it is $b$).
%We define function $\Y{i,j}{m}{a,b}$ as a recurrence, and will use as initial conditions:
%\begin{equation}
%	\Y{-1,j}{m}{X,X}:=
%		\displaystyle
%	  \Z{j,n-1}{m}{X,X}
%\label{eq:Y_in}
%\end{equation}
%The recurrence, as shown below, will increase the interval $[i,j]$ by decreasing $i$ when
%it is not base paired. If it is with a position $k>j$, we increase $j$ to include it.
% Thus, when we need
%to evaluate an interval as $(-1,j)$, all stems between $(0,j)$ are taken into account and the
%structure between $(j,n-1)$ must be a set of independent stems. Therefore,
% all the outside energy between $[j,n-1]$ is
%equal to $\Z{j,n-1}{m}{X,X}$, for any $X\in B$. The recursion itself is as follows.
%\begin{equation}
%	\Y{i,j}{m}{a,b} = \left\{
%  \begin{array}{ll}
%		\displaystyle
%    \sum_{\substack{a'\in \B,\\ \Kron_{a',s_i}\le m}}
%    \Y{i-1,j}{m- \Kron_{a',s_i}}{a',b} &
%    \text{Elif }S_i=-1 \\
%    \displaystyle
%    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_j}\le m}}
%		 e^{\frac{-E_{(i,j),ab \to a'b'}^{\Omega,\beta}}{RT}}\cdot
%    \Y{i-1,j+1}{m- \Kron_{a'b',s_is_j}}{a',b'} &
%   	 \text{Elif }S_{i}=j \land S_{i+1}=j-1\\
%		 \displaystyle
%		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
%		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
%  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
%		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
%     \cdot\Z{j,k-1}{m'}{b,b'} &
%		 \text{Elif }S_{i}=k \geq j\\
%		 \displaystyle
%		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
%		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
%   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
%		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
%     \cdot\Z{k+1,i-1}{m'}{a',b'} &
%		 \text{Elif }-1 < S_{i}=k < i\\
%		 0 & \text{Otherwise}
%  \end{array}\right.
%\label{eq:Y_rec}
%\end{equation}
%The five cases can be broked down as follows.
%\begin{description}
%\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
%as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
%nucleotides $a'$ at position $i$, correcting the number of mutants
%in function of $\Kron_{a',s_i}$.
%\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
%paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
%the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
%to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
%$a',b'\in B^2$ at positions $i$ and $j$ respectively.
%\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
%and is not stacked inside, the 
%only term contributing directly to the energy is the isostericity of the base pair $(i,k)$. 
%We can then consider the outside interval $[i-1,k+1]$ by multiplying it by the the \emph{inside}
%value of the newly included interval (i.e. $[j,k-1]$), for 
%all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
%\item[$-1<S_{i}<i$:]As above but if position $i$ is to pairing with a lower value.
%\item[Else:] In all other cases, we are in a derivation of the SCFG that does not correspond to the 
%secondary structure $S$, and we return $0$.
%
%
%\end{description}
%
%\subsubsection{Combining Inside and Outside values into point-wise mutations probabilities}
%By construction, the partition function over all sequences at exactly $m$ mutations of $s$ can 
%be described in function of the \emph{inside} term as $\Z{0,n-1}{m}{X,X}$,
% for any nucleotide $X\in B$ or
%in function of the \emph{outside} term, for any unpaired position $k$:
%$$
%	\Z{0,n-1}{m}{X,X}
%	\equiv
%	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
%	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a}
%$$
%
%We are now left to compute the probability that a given position is a given nucleotide.
%We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
%Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations. 
%\begin{align}
%	\mathbb{P}(s_i = x\mid s,\Omega, S,M) &:= \frac{\mathcal{W}(i,x,s,\Omega,S,M)}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}\\ 
%\mathcal{W}(i,x,s,\Omega,S,M)&=
% \left\{
%	\begin{array}{ll}
%			\sum_{m=0}^{M}
%			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
%		&\text{If }S_i = -1\\
%			\displaystyle
%			\sum_{m=0}^{M}
%			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
%			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
%     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
%			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
%			\cdot\Z{i+1,k-1}{m'}{x,b}
%		&\text{If }S_i=k>i\\
%    \displaystyle
%			\sum_{m=0}^{M}
%			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
%			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
%     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
%			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
%			\cdot\Z{k+1,i-1}{m'}{b,x}
%		&\text{If }S_i=k<i
%	\end{array}\right.\label{eq:combine}
%\end{align}
%
%In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
%for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
%\begin{description}
%\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
%over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
%of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
%\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
%position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
%base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
%all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
%\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
%\end{description}
%
%\subsection{Complexity considerations}
%Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming. Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation. Furthermore, energy terms $E(\cdot)$ can be accessed in constant time thanks to a simple precomputation (not described)  of the isostericity contributions in $\Theta(n\cdot|\Omega|)$. Computing any given term therefore requires $\Theta(m)$ operations.
%
%In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
%However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
%For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$. Similar constraints hold for the outside computation, resulting in a drastic limitation of the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored. Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ arithmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.