### 2013-05-22 JCB/methods_JCB.tex
Almost there on outside...

%!TEX root = main_RNAPyro_JCB.tex

\section{Methods}
\label{sec:methods}

We introduce a probabilistic model, which aims at capturing both the stability of the folded RNA and its affinity towards a predefined 3D conformation.
To that purpose, a Boltzmann weighted distribution is assumed, based on a pseudo-energy function $\PE{\cdot}$ which includes contributions for both the free-energy and its putative isostericity towards a multiple sequence alignment. In this model, the probability that the nucleotide at a given position needs to be mutated (i.e. corresponds to a sequencing error) can be computed using a variant of the \emph{Inside-Outside algorithm}~\cite{Lari1990} in time which scales linearly with the sequence length.

\subsubsection{Definitions.}
Let $\B:=\left\{\Ab,\Cb,\Gb,\Ub\right\}$ be the set of nucleotides.
Given an RNA sequence $s\in \B^n$, let $s_i$ be the nucleotide at position $i$. Let $\Omega$ be a set of un-gapped RNA sequences of
length $n$. $\Struct$ is a secondary structure without pseudoknots, denoted by a dot-parenthesis string (well-parenthesized expression with dots). In any such expression, matched parentheses induce and unambiguous set of corresponding positions, associated with base-pairing positions, mediated by hydrogen bonds. It follows that if $(i,j)$ and $(k,l)$ are base pairs in $S$, there is no overlapping extremities  $\{i,j\}\cap \{k,l\}=\varnothing$ and either the intersection is empty 
 ($[i,j]\cap[k,l]=\varnothing$) or one is included in the other ($[k,l]\subset[i,j]$ or 
 $[i,j]\subset[k,l])$. Let us finally denote by $\delta: \B^*\times \B^* \to \mathbb{N}^+$ the Hamming distance, i.e. the number of differing positions between two sequences $s'$ and $s''$ such that $|s'|=|s''|$.
\TODOYann{Move above, define secondary structure and add fancy VARNA illustration}

\subsection{Probabilistic Model}
Let $\Omega$ be an gap-free RNA alignment sequence, $S$ its associated secondary structure, 
then any sequence $s$ has probability proportional to its Boltzmann factor
\begin{align*}
  {B}(s) &= e^\frac{-\PE{s}}{RT}, &&\text{with}&\PE{s}&:=\alpha\cdot\ES(s,S)+(1-\alpha)\cdot\EI(s,S,\Omega),
\end{align*}
where $R$ is the Boltzmann constant, $T$ the temperature in Kelvin, $\ES(s)$ and $\EI(s,S,\Omega)$ 
are the free-energy and isostericity contributions respectively (further described below), and $\alpha\in[0,1]$ is an arbitrary parameter that sets the relative weight for both contributions.

\subsubsection{Energy Contribution.}
The free-energy contribution in our pseudo-energy model corresponds to an additive stacking-pairs model, using values from the Turner 2004 model retrieved from the NNDB~\cite{Turner2010}. Given a candidate sequence $s$ for a secondary structure $S$, the free-energy of $S$ on $s$ is given by
\begin{align*}
  \ES(s,S) = \sum_{\substack{(i,j)\to (i',j')\in S\\ \text{stacking pairs}}}\ES^{\beta}_{s_is_j\to s_{i'}s_{j'}} 
\end{align*}
where $\ES^{\beta}_{ab\to a'b'}$ is set to $0$ if $ab=\varnothing$ (no base-pair to stack onto), the tabulated free-energy of stacking pairs $(ab)/(a'b')$ in the Turner model if available, or $\beta\in[0,\infty]$ for non-Watson-Crick/Wobble entries (i.e. neither $\Gb\Ub$, $\Ub\Gb$, $\Cb\Gb$, $\Gb\Cb$, $\Ab\Ub$ nor $\Ub\Ab$). This latter parameter allows to choose whether to simply penalize invalid base pairs, or forbid them altogether ($\beta = \infty$).
The loss of precision due to this simplification of the Turner model remains reasonable since the targeted secondary structure is fixed. For instance, multiloops do not consider base-specific contributions, and therefore their consideration would constitute a criterion for preferring a sequence over another. Furthermore, it greatly eases the design and implementation of dynamic-programming equations. 
\subsubsection{Isostericity Contribution.}
The concept of isostericity score~\cite{Stombaugh2009} is based on the geometric discrepancy (superimposability) of two base-pairs, using individual additive contributions computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Let $s$ be a candidate sequence for a secondary structure $S$, given in the context of a gap-free RNA alignment $\Omega$,  we define the isostericity contribution to the pseudo-energy as
\begin{align*}
  \ES(s,S,\Omega) &= \sum_{\substack{(i,j)\in S\\ \text{pairs}}}\EI^{\Omega}_{(i,j),s_i s_j}, & \text{where}&& 	\EI^{\Omega}_{(i,j),ab}:=
	\frac{
		\sum_{s'\in\Omega}
			\text{\ISO}((s_i',s_j'),(a,b))}
%-		\left(			\text{\ISO}((s_i',s_j'),(s_i,s_j))				\right)	
{		
		|\Omega|
	}
\end{align*}
is the average isostericity of a base-pair in the candidate sequence, compared with the reference alignment.
The $\ISO$ function uses the {Watson-Crick/Watson-Crick} cis isostericity matrix computed by Stombaugh~\emph{et al}~\cite{Stombaugh2009}. Isostericity scores range between $0$ and $9.7$, $0$ corresponding to a perfect isostericity, and a penalty of $10$ is used for missing entries.
The isostericity contribution exponentially favors sequences that are likely to adopt a similar local conformation as the sequences contained in the alignment.
\TODOYann{Avoid duplicate reference to Stombaugh et al}

\subsubsection{Combining contributions.}
Let us remark that any of the individual contributions can be associated to (a subset of) the base-pairs occurring in the structure, possibly complemented, in the case of stacking pairs, with the knowledge of flanking base-pairing nucleotide.
Therefore, let us denote by $\EBP{i,j}{xy}{a'b'}$ the local contribution of a base-pair $(i,j)$ of nucleotides $(a',b')$, surrounded by a stacking pair $(x,y)$ (or $\varnothing$ otherwise), to the pseudo-energy:
\begin{equation}
  \EBP{i,j}{xy}{a'b'}  = \alpha \cdot\ES^\beta_{xy \to a' b'}+(1-\alpha)\cdot\EI^{\Omega}_{(i,j),a'b'}.
\end{equation}

\subsection{Computing the Mutational Profile of Sequences}

Let $s$ be an RNA sequence, $S$ a reference structure, and $m\geq 0$ a desired number of mutations. 
We are interested in  the probability that a given position contains a specific nucleotide, over all sequences having at most $M$ mutations from $s$
%  under the SCFG derivation $S$ 
(formally $\mathbb{P}(s_i = x\mid s,\Omega, \Struct,M)$). 
We define a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990} to compute this
 probability, based on two sets of dynamic programming equations. 
%from the two  functions $\mathcal{Z}_*^*$ and $\mathcal{Y}_*^*$.

The former, defined in Equations~\eqref{eq:Z_in} and~\eqref{eq:Z_rec_A}-\eqref{eq:Z_rec_C}, is analogous to an \emph{inside} computation. Considering an interval of the sequence/secondary structure, it consists in the accumulated contributions of any possible sequences that have suitable Hamming distance within the interval. It is therefore similar to a partition function, i.e. the sum of Boltzmann factors over all sequences within $[i,j]$, 
knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$), within 
$m$ mutations of $s$. The latter, defined by Equations~\eqref{eq:Y_in} and~\eqref{eq:Y_rec},
 computes the \emph{outside} algorithm,   
the partition function over sequences within $m$ mutations of $s$ outside the interval (restricted to two intervals $[0,i]\cup[j,n-1]$), 
knowing  that flanking inner positions $(i+1,j-1)$ contain nucleotides $a$ and $b$ respectively. A suitable multiplicative combination of these terms, computed as shown in Equation~\eqref{eq:combine}, gives the total weight of every possible sequences that support a given base-pair (or unpaired position) and, in turn, the probability of seeing a specific base at a given position.

%Drawing a parallel with stochastic context-free grammars (SCFG), for which the inside-outside algorithm was introduced, the set of sequences can be seen as the language of words having length $n$, generated from a stochastic context-free grammar.
%These rules would limit the number of mutations (i.e. )
% $S$ can be considered as constraining weighing on the shape of parse trees the derivations in a classic 
% of the {SCFG} generating  all  secondary structures of length  $|s|$.

\subsubsection{Inside computation.}
\begin{figure}[t]\centering
\includegraphics[scale=\ScaleDP]{FigDPInsideWrapper}
\caption{Principle of the inside computation (partition function). Any sequence (mutated)  
can be decomposed as a sequence preceded by a, possibly mutated, base 
(Unpaired case), a sequence surrounded by some base-pair (Stacking-pair case), 
or as two sequences segregated by some base-pair (General base-pairing case). In this latter case, mutations must be distributed between sub-sequences.\label{fig:inside}}
\end{figure}

The \emph{Inside} function $\Z{\Struct}{m}{x,y}$ is simply the partition function, i.e. the sum of Boltzmann factors over all sequences for a substructure $\Struct$ (implicitly attached to an interval $[i,j]$ of the sequence), featuring $m$ mutations/errors compared to $s$, and having flanking nucleotides $x$ and $y$. 
Such terms can be computed by recurrence, using the following equation for the initial case:
\begin{equation}
	\forall x,y\in B\times \B, m\in [0,M]:\, \Z{\varepsilon}{m}{x,y}=\left\{
	\begin{array}{ll}
		1 &\text{ If } m = 0\\
		0 &\text{ Otherwise.}
	\end{array}\right.
\label{eq:Z_in}
\end{equation}
In other words, either there is no sequence at distance $m>0$ of the empty sequence, or the only allowed sequence is the empty sequence ($m=0$), having energy $0$. Since the contributions only depend on base pairs, they do not appear in the initial conditions.

The main recursion considers a general structure $\Struct$, flanked by two nucleotides (outside the region of interest) $x$ and $y$, respectively on the $5'$ and $3'$ end of the sequence. As illustrated by Figure~\ref{fig:inside}, it is computed, for each subinterval $[i,j]$, by considering one of the three following cases, dependent on the base-pairing status and context of the leftmost position in the sequence/structure:
\begin{itemize}
\item {\bf Case 1: Unpaired leftmost position.} If the first position  is unpaired in the structure, then $\Struct$ can be further decomposed, using a dot-parenthesis notation, as $\Struct = \ub \Struct'$. Let $a\in B$ be the nucleotide found at the leftmost position in the initial structure, then one has:
\begin{equation}
\label{eq:Z_rec_A}
	\Z{\Struct}{m}{x,y} =
      \sum_{\substack{a'\in \B,\\ \Kron_{a,a'}\le m}}  
      \Z{\Struct'}{m-\Kron_{a,a'}}{a',y}.
\end{equation}
Indeed, any suitable sequence is a concatenation of a, possibly mutated, nucleotide $a'$ at the first position, followed by a sequence over the remaining interval, having $m-\Kron_{a,a'}$ mutations (accounting for a possible mutation at first position), and having flanking nucleotides $a'$ and $y$.
\item {\bf Case 2: Paired ends, stacking onto another base-pair.} If both ends of the considered interval form a base-pair ($\Struct = \op\Struct'\cp$), stacking onto another base pair just outside whole region, then the isosteric contribution of the base-pair must be supplemented with a specific "stacking-pairs" bonus. Let $a$ and $b$ be the nucleotides found on both ends of the interval (positions $i$ and $j$), then one has
\begin{equation}
\label{eq:Z_rec_B}
	\Z{\Struct}{m}{x,y} =
      \sum_{\substack{a',b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
			 e^{\frac{-\EBP{i,j}{xy}{a'b'}}{RT}}
			 \cdot \Z{S'}{m-\Kron_{ab,a'b'}}{a',b'}.
\end{equation}
Any sequence generated here consists of two, possibly mutated, nucleotides $a'$ and $b'$, flanking a sequence over the remaining portion. In order for the total distance to sum to $m$, this portion must feature $m-\Kron_{ab,a'b'}$ additional point wise mutations.

\item {\bf Case 3: Paired leftmost position, but no stacking pairs.} In this case, the structure is split into two parts by the base pair ($\Struct = \op \Struct'\cp\Struct''$). Let us denote by $k$ the partner of position $i$, and by $a$, $b$ and $c$ the bases found at positions $i$, $k$ and $j$ respectively, then one has:
\begin{equation}
\label{eq:Z_rec_C}
	\Z{\Struct}{m}{x,y}=\sum_{\substack{a',b'\in \B^2,\\ \Kron_{ab, a'b'}\le m}}
      \sum_{m'=0}^{m-\Kron_{ab,a'b'}}
   		 e^{\frac{-\EBP{i,k}{\varnothing}{a'b'}}{RT}}
      \cdot\Z{\Struct'}{m-m'-\Kron_{ab,a'b'}}{a',b'}
      \cdot\Z{\Struct''}{m'}{b',y}.
\end{equation}
In other words, if the leftmost position is paired, and the base-pair is not stacking onto another base-pair, then the  only term contributing directly to the energy is the isostericity of the base pair. Admissible sequences for $\Struct$ consist of two paired nucleotides $a'$ and $b'$ at positions $i$ and $k$ respectively, flanking a sequence for $\Struct'$ (over an interval $[i+1,k-1]$), and followed by a (possibly empty) sequence for $\Struct''$ (over $[k+1,j]$). Since the total number of mutations sums to $m$, a parameter $m'$ is introduced to distribute the remaining mutations between the two sequences.
\end{itemize}

\subsubsection{Outside computation.}	
\begin{figure}[t]\centering
\includegraphics[scale=\ScaleDP]{FigDPOutsideWrapper}
\caption{Principle of the outside computation. Note that the outside algorithm uses intermediate results from the inside algorithm, 
therefore its efficient implementation requires a precomputation of the inside contributions.}
\end{figure}

The \emph{Outside} function, $\Y{\Struct}{m}{x,y}$ is the partition function considering only the 
contributions of subsequences excluding a given structure/interval $\Struct$, occupying the open interval $]i,j[$ in the sequence, at Hamming distance exactly $m$ to the initial sequence $s$, and assuming that nucleotides $x$ and $y$ were previously chosen for $i+1$ and $j-1$, the outermost portions of the excluded structure.
\TODOYann{Reformat outside Figure as in Incarnation.}
The associated terms $\Y{\Struct}{m}{x,y}$ can then be computed recursively, considering as the base case any prefix $\Struct$ of the complete structure:
\begin{equation}
	\forall x,y\in B\times \B, \forall \Struct'\text{ s.t. }\Struct=\Struct'.\Struct'', m\in [0,M]:\, \Y{\Struct^\ast}{m}{x,y}=\Z{\Struct''}{m}{y,X}
\label{eq:Y_in}
\end{equation}
where $X\in\B$ can be any nucleotide, and will not affect further computations.

The main recurrence below extends the excluded structure leftward, and considers four cases depending on its base-pairing status and directionality:
\begin{itemize}
\item {\bf Case 1: Unpaired position.}
\begin{equation}
	\Y{\Struct}{m}{x,y} = \sum_{\substack{a'\in \B,\\ \Kron_{a,a'}\le m}}
    \Y{\ub\Struct}{m- \Kron_{a,a'}}{a',y}
\label{eq:Y_rec_A}
\end{equation}
\item {\bf Case 2: Stacking base-pair.}
\begin{equation}
	\Y{\Struct}{m}{x,y} = 
    \sum_{\substack{a'b'\in \B^2,\\ \Kron_{ab,a'b'}\le m}}
		 e^{\frac{-\EBP{i,j}{xy}{a'b'}}{RT}}\cdot
    \Y{\op\Struct\cp}{m- \Kron_{ab,a'b'}}{a',b'} 
\label{eq:Y_rec}
\end{equation}
\item {\bf Case 3: Next position paired rightward.}
\begin{equation}
	\Y{i,j}{m}{a,b} = \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_is_k}\le m}}
		 \sum_{m'=0}^{m-\Kron_{a'b',s_is_k}}
  		 e^{\frac{-E_{(i,k),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
		 \cdot\Y{i-1,k+1}{m- \Kron_{a'b',s_is_k} - m'}{a',b'}
     \cdot\Z{j,k-1}{m'}{b,b'} 
\label{eq:Y_rec}
\end{equation}
\item {\bf Case 4: Next position paired leftward.}
\begin{equation}
	\Y{i,j}{m}{a,b} = 
		 \sum_{\substack{a'b'\in \B^2,\\ \Kron_{a'b',s_ks_i}\le m}}
		 \sum_{m'=0}^{m-\Kron_{a'b',s_ks_i}}
   	 e^{\frac{-E_{(k,i),\varnothing\to a'b'}^{\Omega,\beta}}{RT}}
		 \cdot\Y{k-1,j}{m- \Kron_{a'b',s_ks_i} - m'}{a',b}
     \cdot\Z{k+1,i-1}{m'}{a',b'}
\label{eq:Y_rec}
\end{equation}
\end{itemize}
\TODOYann{Rephrase in term of tree, as in Incarnation??}
The five cases can be broked down as follows.
\begin{description}
\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, then the value is the same
as if we decrease the lower interval bound by $1$ (i.e. $i-1$), and consider all possible
nucleotides $a'$ at position $i$, correcting the number of mutants
in function of $\Kron_{a',s_i}$.
\item[$S_{i}=j$ and $S_{i+1}=j-1$:] If nucleotide $i$ is paired with $j$ and nucleotide $i+1$ is
paired with $j-11$, we are in the only case were stacked base pairs can occur. We thus add
the energy of the stacking and of the isostericity of the base pair $(i,j)$. What is left
to compute is the \emph{outside} value for the interval $[i-1,j+1]$ over all possible nucleotides 
$a',b'\in B^2$ at positions $i$ and $j$ respectively.
\item[$S_{i}=k \geq j$:]If nucleotide $i$ is paired with position $k\geq j$, 
and is not stacked inside, the 
only term contributing directly to the energy is the isostericity of base pair $(i,k)$. 
Therefore, we consider the outside interval $[i-1,k+1]$, multiplying it by the \emph{inside}
value of the newly included interval (i.e. $[j,k-1]$), for 
all possible values $a',b'\in B^2$ for nucleotides at positions $i$ and $k$ respectively.
\item[$-1<S_{i}<i$:]As above but if position $i$ is paired with a lower value.
\item[Else:] Any other derivation of the SCFG does not correspond to the 
secondary structure $S$, and we return $0$.

\end{description}

\subsubsection{Combining Inside and Outside Values into Point-Wise Mutations Probabilities.}
By construction, the partition function over all sequences at exactly $m$ mutations of a reference sequence $s$ can 
be either described from the \emph{inside} contribution $\Z{0,n-1}{m}{X,X}$ of the whole sequence,
$\forall X\in B$, or from \emph{outside} terms as:
$$
	\Z{0,n-1}{m}{X,X}
	\equiv
	\sum_{\substack{a\in \B,\\ \Kron_{a,s[k]}\le m}}	
	\Y{k-1,k+1}{m-\Kron_{a,s[k]}}{a,a},\forall k \text{	unpaired.}
$$

We are now left to compute the probability that a given position is a given nucleotide.
We leverage the \emph{Inside-Outside} construction to immediately obtain the following $3$ cases.
Given $i\in[0,n-1],x\in B$, and $M\geq 0$ a bound on the number of allowed mutations, one defines
\begin{equation}
	\mathbb{P}(s_i = x\;|\; M) := \frac{\mathcal{W}^M_{\substack{i, [x]}}}{\sum_{m=0}^{M}\Z{0,n-1}{m}{X,X}}\label{eq:normalize}
\end{equation}
where $\mathcal{W}{\substack{\ast\\ \ast}}$ is defined by:
\begin{equation}
\resizebox{\textwidth}{!}{%
$ \mathcal{W}^M_{\substack{i, [x]}} =  \left\{
	\begin{array}{ll}
			\sum_{m=0}^{M}
			\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}
		&\text{If }S_i = -1\\
			\displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{xb,s_is_k\leq m}}}
			\sum_{m'=0}^{m-\Kron_{xb,s_is_k}}
     	 e^{\frac{-E_{(i,k),\varnothing\to xb}^{\Omega,\beta} }{RT}}
			\cdot\Y{i-1,k+1}{m-\Kron_{xb,s_is_k-m'}}{x,b}
			\cdot\Z{i+1,k-1}{m'}{x,b}
		&\text{If }S_i=k>i\\
    \displaystyle
			\sum_{m=0}^{M}
			\sum_{\substack{b\in B\\\Kron_{bx,s_ks_i\leq m}}}
			\sum_{m'=0}^{m-\Kron_{bx,s_ks_i}}
     	 e^{\frac{-E^{\Omega, \beta}_{(k,i),\varnothing\to bx}}{RT}}
			\cdot\Y{k-1,i+1}{m-\Kron_{bx,s_ks_i-m'}}{b,x}
			\cdot\Z{k+1,i-1}{m'}{b,x}
		&\text{If }S_i=k<i
	\end{array}\right.
$}\label{eq:combine}
\end{equation}

\TODOYann{Rephrase as tree constructs}
\TODOVlad{Add illustration of Inside/Outside combination (adapted from RECOMB presentation??)}
In every case, the denominator is the sum of the partitions function of exactly $m$ mutations, 
for $m$ smaller or equal to our target $M$. The numerators are divided in the following three cases.
\begin{description}
\item[$S_i=-1$:] If the nucleotide at position $i$ is not paired, we are concerned by the weights
over all sequences which have at position $i$ nucleotide $x$, which is exactly the sum
of the values of $\Y{i-1,i+1}{m-\Kron_{x,s_i}}{x,x}$, for all $m$ between $0$ and $M$.
\item[$S_i=k>i$:] Since we need to respect the derivation of the secondary structure $S$, if 
position $i$ is paired, we must consider the two partition functions. The \emph{outside} of the 
base pair, and the \emph{inside}, for all possible values for the nucleotide at position $k$, and
all possible distribution of the mutant positions between the inside and outside of the base pair. We also add the term of isostericity for this specific base pair.
\item[$S_i=k<i$:] Same as above, but with position $i$ pairing with a lower position.
\end{description}

\subsection{Complexity Considerations}
Equations~\eqref{eq:Z_rec} and~\eqref{eq:Y_rec} can be computed using dynamic programming. Namely, the $\mathcal{Z}^{*}_{*}$ and $\mathcal{Y}^{*}_{*}$ terms are computed starting from smaller values of $m$ and interval lengths, memorizing the results as they become available to ensure constant-time access during later stages of the computation. Furthermore, energy terms $E(\cdot)$ can be accessed in constant time thanks to a simple precomputation (not described)  of the isostericity contributions in $\Theta(n\cdot|\Omega|)$. Computing any given term therefore requires $\Theta(m)$ operations.

In principle, $\Theta(m\cdot n^2)$ terms, identified by $(m,i,j)$ triplets, should be computed.
However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of intervals $(i,j)$.
For instance, the inside algorithm only requires computing intervals $[i,j]$ that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$. Similar constraints hold for the outside computation, resulting in a drastic limitation of the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored. Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ arithmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.