### 2013-05-26 JCB/conclusion_JCB.tex

%!TEX root = main_RNAPyro_JCB.tex
\section{Conclusion}
\label{sec:conclusion}

\TODOJerome{Etoffer la conclusion}
In this article we presented a new and efficient way of exploring the mutational landscape of an RNA under structural constraints,
and apply our techniques to identify and fix sequencing errors. In addition, we introduced a new scoring scheme to measure the
likelihood of sequencing errors that combines the classical nearest-neighbour energy model parameters \cite{Turner2010} to the recently introduced
isostericity matrices \cite{Stombaugh2009}. The latter accounting for geometrical discrepancies occurring during base pair replacements.

We combined our algorithm for exploring the mutational neighbourhood of an input sequence with known secondary structure to
this new pseudo energy model, and create a tool to predict point-wise sequencing errors in structured RNAs.
Importantly, our algorithm runs in  $\Theta(n\cdot(|\Omega|+M^2))$ time and $\Theta(n\cdot(|\Omega|+M))$ memory, where $n$ is the length of the RNA,
$M$ the number of mutations and $\Omega$ the size of the multiple sequence alignment. This achievement enables us to envision applications to high-throughput sequencing pipe-lines.

We validated our model on the 5s rRNA and 16s rRNA (See Sec.~\ref{sec:results}) and showed that our technology enables us to recover 
mutational errors with [INSERT PERFS]. Interestingly, we observed that using the isostericity matrices alone yields higher performance than with the 
nearest-neighbour energy model alone. This findings supports our hypothesis that isostericity matrices provide a valuable source of informations that
can be efficiently used for RNA sequence and structure analysis.

Our techniques are designed to correcting point-wise errors in structured regions (i.e. base paired nucleotides). Nonetheless, our software 
\texttt{RNApyro} can be easily combined with other methodologies previously developed for correcting other types of sequencing errors such as indels in 
unstructured regions or repeats \cite{Quinlan2008,Quince:2009uq}. In further works, we also intend to include in our model errors stemming from insertions or deletions. It is indeed theoretically possible to considers these scenario within our dynamic programming scheme \cite{Waldispuhl:2002fk} with minor impacts of the complexity.

Finally, we hope that integrating our software to current sequencing pipe-lines used in metagenomics studies will permit to improve the estimate of microbial diversity.