### 2013-07-01 JCB/acknowledgments_JCB.tex
update text commnet ROC

%!TEX root = main_RNAPyro_JCB.tex
\section{Acknowledgments}
\label{sec:acknowledgments}
The authors would like to thank Rob Knight for his suggestions and comments.
This work was funded by the French Agence Nationale de la Recherche (ANR) through the {\sc Magnum} {\tt ANR 2010 BLAN 0204} 
project (to YP), the FQRNT team grant 232983 and NSERC Discovery grant 219671 (to JW).
This paper has been developed as a result of a mobility stay funded by the
Erasmus Mundus Programme of the European Commission under
the Transatlantic Partnership for Excellence in Engineering TEE
Project (VR).